### 2009-09-16 latex/paper1/contents.tex
paper

Software modularization recovery algorithms automatically recognize a system's
modular structure by analyzing its implementation.
Due to the lack of well document software systems, though, the issue of testing
these algorithms is still underexplored, limiting both their adoption in the
industry and the development of better algorithms.
We propose to rely on software models to produce arbitrarily large test sets. In
this paper we consider three such models and analyze how similar the artifacts
they produce are from artifacts from real software systems.
\cite{Pollet2007}

@@

Development of large-scale software systems is a challenge.

@@

POR ISSO SURGIRAM software modularization recovery algorithms, also known as
software clustering algorithms or software architecture recovery algorithms. In
its most common flavor, these algorithms analyze the dependencies between
implementation components, such as classes, and then group them into modules
such as there are few dependencies between classes in distinct modules.

@@

\section{Related Work}

\section{Software Modularization Recovery: an Overview}

Aka software architecture recovery, software architecture reconstruction,
software clustering...

It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...

DEFINITIONS

external edge...

%\section{Software Networks and Software Clustering}

directed graph, (un)weighted

\section{Network Theory}

Network theory research studies general properties of many types of networks by
using statistical analysis. In the last decade, it has been found that many
networks arising from sociology, biology, technology and other domains present
remarkable structural similarities. It has been shown that, in these networks,
the number of vertices connected to k edges, N(k), is proportional to
$k^{\-gamma}$, where $\gamma$ is a positive constant. These networks were called
scale-free networks.

@@

(in fact we can't argue the degree distribution is perfectly fit by a power law.
Anyway, the distribution is much more assymetric than normal ou Poisson's)

software indegree distribution (power law), outdegree distribution (not so power
law)

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to generate networks that are scale-free. In this section we present three
models that generate directed networks with built-in modular decomposition:
BCR+, CGW and LR. The first two models grow networks using a mechanism called
preferential attachment CITE Barabasi, by which nodes with many edges tend to
receive more edges in the process.

@@

\subsection{BCR plus}

The BCR model aims to model the network of hyperlinks between web pages as a
directed graph without modules \cite{Bollobas2003}. It was proven analitically
to generate scale-free networks. We have developed an extension to this model,
which we will call BCR+, that adds the concept of module. The model accepts the
following parameters:

\begin{itemize}
\item number of vertices, $n$;
\item a graph, $G$;
\item three probabilities, $p$, $q$, and $r$, such as $p + q + r = 1$;
\item a probability, $mu$;
\item two real numbers, din and dout.
\end{itemize}

In a network with modules, one can define a MDG as a graph where the two
following conditions hold:
* each vertex represents a module in the original network;
* there is an edge from M1 to M2 in the MDG only if there is an edge from 
v1 to v2 in the original network, where v1 in M1 and v2 in M2.

The BCR+ model generates networks whose MDG is equal to G by adding vertices and
edges to an initial network until it reaches n vertices. The initial network is
isomorphic to G: it contains one vertex for each module and one edge for each
edge in G, connecting vertices whose corresponding modules are are connected.
Thus, the networks's MDG is equal to G from the very beginning.

After that, the algorithm consists of successive applications of one out of
three operations on the network: (1) adding a vertex with an outgoing edge; (2)
adding a vertex with an ingoing edge; (3) adding a edge between existing
vertices. The choice of the vertices that will be connected by a new edge,
although non-deterministic, is not fully random. The probability that a
particular vertex v is choosen, P(v), is proportional to a function of the
in-degree or of the out-degree of the node. We say that we choose a vertex
within a set V according to a function f if

  P(choose v) = f(v) / sum(v in V, f(v))

@@

Before we present the algorithm in detail, using pseudo-code, we must explain
some definitions. V is the set of vertices in the network being generated. This
set grows as the algorithm is executed. Being v a vertex, we define the
following three functions:

* out-neighbors(v): the set of all vertices that are connected to v by an edge
starting in v.
* same-module(v): the set of all vertices that are in the same module as v (except
v itself).
* other\_modules(v): the set of all vertices that are in modules that are connected
to v's module (in the graph G) by an edge starting in v's module.

@@

\begin{figure*}
\begin{verbatim}
while the network has less than n vertices:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex v to w's module
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex v to w's module
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	choose a vertex w within (other\_modules(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within (same-module(v) - neighbors(v)) according to f(x) = din + in-degree(x)
      add an edge from v to w
  end
\end{verbatim}
\end{figure*}

We can see that the probabilities p, q and r control how often each operation is
performed. Because the operation associated with probability r does not add any
vertices, greater values of r imply more edges in the resulting network. It is
easy to see that nodes connected to many ingoing edges are more likely to get
another ingoing edge, and the same reasoning can be applied to outgoing edges.
The parameter din can alleviate the handicap by providing a ``base in-degree''
that is applied to all vertices when computing the probabilities. Consider
two vertices, v1 with in-degree 4, and v2 with in-degree 8. If din = 0, v2 is
twice more likely to receive a new incoming edge; if, otherwise, din = 4, v2 is
only 3/2 more likely to receive the edge.

The parameter mu controls the proportion of edges between vertices in distinct
modules. Lower values of mu lend to networks that are more modular.

The BCR+ model is a growth model, meaning that the network is generated vertex
by vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down design methodologies CITE.

@@

The CGW model was proposed to model the evolution of software systems organized in
modules. It was proven formally to generate scale-free networks. \cite{Chen2008}
It accepts the following parameters:

* n, m
* Probabilities p1, p2, p3, p4, summing 1
* Natural numbers e1, e2, e3, e4
* alpha

Just like BCR+, this is a growth model. Its initial network is composed of two
vertices belonging to the same module and a directed edge between them. The
remaining m - 1 modules are initially empty. Because the original implementation
of the model is not available, we describe in detail the algorithm we
implemented:

\begin{figure*}
\begin{verbatim}
while the network has less than n vertices:
  choose operation (p1, p2, p3, p4)
  operation 1: (adding a vertex with e1 edges)
    create a new vertex v and add it to a randomly chosen module
    do e1 times:
      choose a vertex w according to mbpa(v)
      add an edge from v to w
  operation 2: (adding e2 edges)
    do e2 times:
      choose a vertex v randomly
      choose a vertex w according to mbpa(v)
      add edge from v to w
  operation 3: (rewiring e3 edges)
    do e3 times:
      choose a vertex v randomly
      choose an edge e randomly within edges that start in v
      choose w according to mbpa(v)
      remove e
      add an edge from v to w
  operation 3: (removing e4 edges)
    do e4 times:
      remove and randomly chosen edge
\end{verbatim}
\end{figure*}

@@

The LF model is a very flexible model that can generated weighted directed
networks with overlapping modules, that is, in which a vertex can belong to more
than one module. Unlike the previous models, this is not a growth model: all
vertices are generated at once and then the edges are added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at http://.

@@

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for three distinct networks.

Following the work by Milo et al. \cite{Milo2002}, similarity between two
networks can be measured by computing Pearson's correlation coefficient between
the corresponding TCPs, which yields a value between -1 (most dissimilar) and 1
(most similar):

@@

\begin{figure*}[!t]
\center
\subfigure[ref1][Software
network]{\includegraphics[width=2.5in]{triads-software}}
\hfil
\subfigure[ref2][Non-software network]{\includegraphics[width=2.5in]{triads-nonsw}%
}
\caption{Triad profiles}
\label{fig:profiles}
\end{figure*}

\subsection{Data Set}

@@

Then we measured the S-score for each non-software network. The majority of the
networks (X\%) have a S-score lower than 0.83, which shows that they
are ... Networks extracted from the social network Facebook have negative
S-score.

A few networks, though, showed high S-score. The network
polblogs, showing links between blogs on politics, had similarity 0.97. The
neural network of C Elegans also had a high similarity (0.88). Further
investigation is needed in order to discover why does it happen and whether
auxiliary metrics can differentiate these networks from software networks.

\subsection{A Network Classification Model}

@@

As we have shown on the previous section, there are non-software networks with
high S-scores, hence it is impossible to build a perfect classification model,
regardless of the threshold. Nonetheless, such a model can be evaluated by its
precision and recall. Consider a data set with both software and non-software
networks. Let $S$ be the set of all software networks, and $L$ the set of all
networks that were classified as software-like. The precision of the model is

$$
precision: \frac{S \cap L}{L}
$$

@@

$$
recall: \frac{S \cap L}{S}
$$

Increasing the threshold has the effect of reducing the recall, because fewer
software networks are classified as software-like. Decreasing the threshold has
the effect of reducing the precision, since many non-software networks are
classified as software-like.

The choice of a proper threshold, thus, depends on whether it is more important
to have high precision of high recall. Because our research hypothesis is that
networks synthesized by the presented models are software-like, higher precision
means a stronger test, as fewer networks are classified as software-like.

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be X\%, which is too low, meaning the most software networks are
misclassified. So we chose the value 0.88, that is immediatly greater than the second greater
non-software network S-score. With this value, we have a high recall (X\%) and a
still high precision (X\%).

@@

We have chosen five different module dependency networks, which where extracted
from actual dependencies between archives of five different software systems of
our sample. The module dependency networks are shown on Figure
\ref{fig:architectures}.

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{architectures}
\caption{Module dependency graphs}
\label{fig:architectures}
\end{figure*}

%02-gef
%04-ibatis
%08-megamek
%16-findbugs
%32-zk

The probabilities p, q and r were given all possible values from 0.0 up to 1.0,
in 0.2 steps, such as the sum of the probabilities was 1. Since the only events
that create vertices are those associated with probabilities p and q, we imposed
the additional restriction that p + q > 0.

For deltain and deltaout we assigned the integer numbers from 0 to 4. TODO: why

Finally, we chose mu from 0.0 to 0.6 in 0.2 steps. It does not make sense to
choose higher values since they mean that there will be more edges connecting
different modules.

In total, 9,500 networks were synthesized.

@@

Like BCR, we choose mu ranging from 0.0 to 0.6 in 0.2 steps. For the remaining
parameters we selected values from our sample of software systems. For degexp,
..., we picked the minimum, the median and the max values.

Total: 1,296 networks

We have varied each of the probabilities p1, p2, p3, and p4 from 0.0 to 1.0, in
0.2 steps, with $p1 > 0$ and $p1 + p2 + p3 + p4 = 1$. Each of the parameters
$e1$, $e2$, $e3$, and $e4$ assumed the values 1, 2, 4, and 8.

% We have imposed
%the additional restriction that $2 \cdot p4\cdot e4 \ge p1 \cdot e1 + p2\cdot e2$, 
%so the number of edges created is at least twice the number of edges removed
%(otherwise 
% TODO: DOES NOT MAKE SENSE

alpha in -1, 0, 1, 10, 100, 1000. 
m in 2, 4, 8, 16, 32 (just like bcr and lfr)

In total, 38,790 networks were synthesized.

@@

We used our classifier ...

Table: Model | Number of Networks | Realistic Networks | Percent \%

Show some graphs: histogram of average correlations for each model.
\ref{fig:histograms}

\begin{figure*}[!t]
\center

\subfigure[Software network]{\includegraphics[width=1.0in]{hist-cgw}}

\subfigure[Non-software network]{\includegraphics[width=1.0in]{hist-bcr}}

\subfigure[Non-software network]{\includegraphics[width=1.0in]{hist-lf}}

\caption{Triad profiles}
\label{fig:histograms}
\end{figure*}

All models produce networks that resemble software networks.  For some
parameters, though, the networks are not realistic.

\subsection{Parameter-Based Classification Models}

1R

Naive Bayes

\subsection{Homogeinity}

Pick realistic networks from a model. Are they similar to each other? (see
standard deviation) In other words: do the parameters make a difference?

Are they similar to networks generated by the other models? In other words: are
the models equivalent?

\section{Threats to validity}

% External validity (EV): can it be generalized?

%Structural information isn't enough for a expert do produce a
%decomposition (s/he may use data such as names and external documentation). 
%
%Even when considering only structural information, is it true
%that experts would find a decomposition similar to the reference decomposition
%imposed by the model?

We generated only one network for each set of parameters. (as we've shown, some
parameters are redundant as they do not change significantly the realism)

Some clustering algorithms use weights and they weren't studied here.

EV: We've only studied 65 systems, which is not that much.

We only studied object-oriented systems implemented in Java. Maybe the results
would be different if we studied systems implement in other languagens or using
other paradigms. The choice of a particular technique for extracting
dependencies (static analysis) may also have impact on the structure of the
networks.

% Koschke says that experts decompositions vary by no more than 80? percent.

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously studied with real networks
\cite{Wu2005}. After that we will be able to compare the results obtained by
both approaches.

\section{Appendix A: list of networks}

TODO: Show as Table: Name, Vertices, Edges

Software networks:

>>>>

Software modularization recovery algorithms automatically recognize a system's
modular structure by analyzing its implementation.
Due to the lack of well document software systems, though, the issue of testing
these algorithms is still underexplored, limiting both their adoption in the
industry and the development of better algorithms.
We propose to rely on software models to produce arbitrarily large test sets. In
this paper we consider three such models and analyze how similar the artifacts
they produce are from artifacts from real software systems.

@@

Development of large-scale software systems is a challenge. \cite{Northrop2006}

@@

POR ISSO SURGIRAM software modularization recovery algorithms, also known as
software clustering algorithms or software architecture recovery algorithms. In
its most common flavor, these algorithms analyze the dependencies between
implementation components, such as classes, and then group them into modules
such as there are few dependencies between classes in distinct modules.
\cite{Pollet2007}

TODO: sw clustering is a task of reverse engineering \cite{Tonella2007}.

@@

%\section{Related Work}
%
%\section{Software Modularization Recovery: an Overview}
%Aka software architecture recovery, software architecture reconstruction,
%software clustering...
%It's an task of reverse engineering, as defined by Tonella \cite{Tonella2007}...
%DEFINITIONS
%external edge...

%\section{Software Networks and Software Clustering}
%
%directed graph, (un)weighted

\section{Software Networks}

Define: Component. Component dependency network. Directed unweighted graph.

Clustering based on dependencies.

Network theory research studies general properties of many types of networks by
using statistical analysis. In the last decade, it has been found that many
networks arising from sociology, biology, technology and other domains present
remarkable structural similarities. It has been shown that, in these networks,
the degree distribution is a power law: the number of vertices connected to k
edges, N(k), is proportional to $k^{\-gamma}$, where $\gamma$ is a positive
constant. These networks are called scale-free networks.

@@

%(in fact we can't argue the degree distribution is perfectly fit by a power law.
%Anyway, the distribution is much more assymetric than normal ou Poisson's)
%
%software indegree distribution (power law), outdegree distribution (not so power
%law)

@@

Many models were proposed to explain the formation of scale-free networks. These
models are simple algorithms that can be proven, either formally or empirically,
to synthesize networks that are scale-free. Most models, though, do not account
for modular structure, a feature present in some networks, in particular
software networks. In this section we present three models that generate
directed networks with built-in modular decomposition: BCR+, CGW and LR.

%The first two models grow networks using a mechanism called
%preferential attachment CITE Barabasi, by which nodes with many edges tend to
%receive more edges in the process.

@@

\subsection{BCR+}

The BCR model \cite{Bollobas2003} aims to model the network of hyperlinks
between web pages as a directed graph without modules. It was proven
analitically to generate scale-free networks. We have developed an extension to
this model, called BCR+, that adds modules to the construction of the network.
The model accepts the following parameters:

\begin{itemize}
\item number of vertices, $n$;
\item a graph, $G$;
\item three probabilities, $p$, $q$, and $r$, such as $p + q + r = 1$;
\item a probability, $mu$;
\item two real numbers, $din$ and $dout$.
\end{itemize}

In a network with modules, one can define a module dependency graph (MDG) as a
graph where the two following conditions hold:

\begin{itemize}
\item each vertex represents a module in the original network; 
\item there is an edge from $M_1$ to $M_2$ in the MDG only if there is an edge
from $v_1$ to $v_2$ in the original network, where $v_1 \in M_1$ and $v_2 \in
M_2$.
\end{itemize}

The BCR+ model generates networks whose MDG is equal to $G$ by adding vertices
and edges to an initial network until it reaches $n$ vertices. The initial
network is isomorphic to $G$: it contains one vertex for each module and one
edge for each edge in $G$, connecting vertices whose corresponding modules are
connected. Thus, the MDG for the generated network is equal to $G$ from the very
beginning.

After that, the algorithm consists of successive applications of one out of
three operations on the network: (1) adding a vertex with an outgoing edge; (2)
adding a vertex with an ingoing edge; (3) adding a edge between existing
vertices. In operation 3, the edge can connect either vertices that belong
to the same module or vertices in distinct modules.

The choice of the vertices that will be connected by a new edge, although
non-deterministic, is not fully random. The probability that a particular vertex
v is choosen, P($v$), is proportional to a function of the in-degree or the
out-degree of the node. We say that we choose a vertex within a set of vertices
$S$ according to a function f if

$$
  P(v) ~=~ \frac{ f(v) }{ \sum_{v \in S, f(v)} }.
$$

@@

Before we present the algorithm in detail, using pseudo-code, we must define
some functions:

%Let V be the set of vertices in the network being generated. This
%set grows as the algorithm is executed. Being v a vertex, we define the
%following three functions:

\begin{itemize}

\item in-degree(v): the number of edges that enter in the vertex v;

\item out-degree(v): the number of edges that leave the vertex v;

\item module(v): the module to which the vertex v belongs;

\item out-neighbors(v): the set of all vertices that are connected to the vertex
v by an edge starting in v.

\item same-module(v): the set of all vertices that are in the same module as the
vertex v, except for v itself and vertices in out-neighbors(v).

\item other\_modules(v): the set of all vertices that are in modules that are
connected to v's module (in the graph G) by an edge starting in v's module,
except for vertices in out-neighbors(v).

\end{itemize}

@@

\begin{figure*}
\begin{verbatim}
while the number of nodes in the network is less than n:
  choose one of the operations according to probabilities (p, q, r)
  operation 1: (add a vertex with an outgoing edge)
    choose a vertex w within V according to f(x) = din + in-degree(x)
    add a new vertex, v, to module(w)
    add an edge from v to w
  operation 2: (add a vertex with an ingoing edge)
    choose a vertex w within V according to f(x) = dout + out-degree(x)
    add a new vertex, v, to module(w)
    add an edge from w to v
  operation 3: (add an edge between two existing vertices)
    choose a vertex v within V according to f(x) = dout + out-degree(x)
    choose a case according to probabilities (mu, 1 - mu)
    case 1: (distinct modules)
  	  choose a vertex w within other\_modules(v) according to f(x) = din + in-degree(x)
      add an edge from v to w
    case 2: (same module)
      choose w within same-module(v) according to f(x) = din + in-degree(x)
      add an edge from v to w
\end{verbatim}
\end{figure*}

We can see that the probabilities $p$, $q$, and $r$ control how often each
operation is performed. Because the operation associated with probability $r$
does not add any vertices, greater values of $r$ imply more edges in the
resulting network. It is easy to see that nodes connected to many ingoing edges
are more likely to get another ingoing edge, and the same reasoning can be
applied to outgoing edges. The parameter $din$ can alleviate the handicap by
providing a ``base in-degree'' that is applied to all vertices when computing
the probabilities. Consider two vertices, $v_1$ with in-degree 4, and $v_2$ with
in-degree 8. If $din = 0$, $v_2$ is twice more likely to receive a new incoming
edge; if, otherwise, $din = 4$, $v_2$ is only $\frac{3}{2}$ more likely to
receive the edge.

The parameter $mu$ controls the proportion of edges between vertices in distinct
modules. Lower values of $mu$ lead to networks with weakly coupled modules.

The BCR+ model is a growth model, meaning that the network is generated vertex
by vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down system design methodologies CITE.

@@

The CGW model \cite{Chen2008} was proposed to model the evolution of software
systems organized in modules. It was proven formally to generate scale-free
networks.  It accepts the following parameters:

\begin{itemize}
\item Number of vertices, n;
\item Number of modules, m;
\item Probabilities p1, p2, p3, p4, summing 1;
\item Natural numbers e1, e2, e3, e4;
\item Number alpha, with $alpha \ge -1$
\end{itemize}

Just like BCR+, this is a growth model. Its initial network is composed of two
vertices belonging to the same module and a directed edge between them. The
remaining $m - 1$ modules are initially empty.

After that, the algorithm consists of successive applications of one out of four
operations on the network, which are chosen according to probabilities p1, p2,
p3, and p4: 
(1) adding a vertex with e1 outgoing edges; 
(2) adding e2 edges;
(3) rewiring e3 edges;
(4) removing e4 randomly chosen edges.

Whenever an edge is added from a vertex $v$ to a vertex $w$ (operations 1, 2,
and 3), $v$ is chosen randomly, and $w$ is chosen according to the following
probability function:

$$
  P_v(w) ~=~ \frac{ f_v(w) }{ \sum_{w \in S, f_v(w)} },
$$

where

$$
f_v(w) = 1 + degree(w) \cdot (1 + \alpha), if module(w) = module(v)
, 1 + degree(w), otherwise
$$

Because the original paper did not provide an implementation for the model, we
now present the pseudo-code for our implementation:

\begin{figure*}
\begin{verbatim}
while the network has less than n vertices:
  choose operation (p1, p2, p3, p4)
  operation 1: (adding a vertex with e1 outgoing edges)
    create a new vertex v and add it to a randomly chosen module
    do e1 times:
      choose a vertex w according to f_v(w)
      add an edge from v to w
  operation 2: (adding e2 edges)
    do e2 times:
      choose a vertex v randomly
      choose a vertex w according to f_v(w)
      add edge from v to w
  operation 3: (rewiring e3 edges)
    do e3 times:
      choose a vertex v randomly
      choose an edge e randomly within edges that start in v
      choose w according to f_v(w)
      remove e
      add an edge from v to w
  operation 3: (removing e4 edges)
    do e4 times:
      remove and randomly chosen edge
\end{verbatim}
\end{figure*}

@@

The LF model \cite{Lancichinetti2009} is a very flexible model that can
generated weighted directed networks with overlapping modules, that is, in which
a vertex can belong to more than one module. Unlike the previous models, this is
not a growth model: all vertices are generated at once and then the edges are
added.

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the following
parameters:

\begin{itemize}
\item number of nodes, $n$;
\item average in-degree, $k$, with $k < n$;
\item maximum in-degree, $max_k$, with $max_k \ge k$ and $max_k < n$;
\item mixing parameter, $\mu$, with $0 \le \mu \le 1$;
\item minus exponent for the degree distribution, $\gamma$;
\item minus exponent for the module size distribution, $\beta$;
\item minimum module size, $min_m$;
\item maximum module size, $max_m$, with $max_k \ge min_k$;
\end{itemize}

The sizes of the modules are selected from a power law distribution with
exponent $-\beta$. The mixing parameter, $\mu$ is the proportion of edges in the
network that connect vertices from distinct modules. In this model, some
combinations of parameter values are unfeasible. For example, if $n = 100$ then
$min_m$ cannot be 60 (otherwise there would be modules smaller than the minimum
module size).

@@

By counting how many times each triad appears in a network, one can build a
triad concentration profile (TCP), which is a vector with 13 numbers that
summarize the local structure of the network. Figure \ref{fig:profiles} shows
the TCP for networks from distinct domains.

Following the work by Milo et al. \cite{Milo2004}, similarity between two
networks can be measured by computing Pearson's correlation coefficient between
the corresponding TCPs, which yields a value between -1 (most dissimilar) and 1
(most similar):

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see Appendix A). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Data Set}

@@

Then we measured the S-score for each non-software network. The majority of the
networks (X\%) had a S-score lower than 0.83, which is the lowest S-score for
software networks in the sample. Some networks, showing friendship between
students, showed negative S-score, meaning that they are very different from
software networks.

Two networks, though, showed high S-score, above 0.83. The network of links
between blogs on politics showed S-score 0.97. The neural network of the worm
C. Elegans also showed a high S-score (0.88). Further investigation is needed
in order to discover the reasons behind these high S-scores and whether
auxiliary metrics can differentiate these networks from software networks.

\subsection{A Network Classification Model} \label{sec:classmodel}

@@

As we have shown on the previous section, there are non-software networks with
high S-scores, hence it is impossible to build a perfect classification model,
regardless of the threshold. Nonetheless, such a model can be evaluated by its
precision and recall. Consider a data set with both software and non-software
networks. Let $S$ be the set of all software networks, and $L$ the set of all
networks that were classified by the model as software-like. The precision of
the model is

$$
\mathrm{precision}: ~\frac{S \cap L}{L},
$$

@@

$$
\mathrm{recall}: ~\frac{S \cap L}{S}.
$$

Increasing the threshold has the effect of reducing the recall, because fewer
software networks are classified as software-like. Decreasing the threshold has
the effect of reducing the precision, since more non-software networks are
classified as software-like.

The choice of a proper threshold, thus, depends on whether it is more important
to have high precision or high recall. Because our research hypothesis is that
networks synthesized by the presented models are software-like, higher precision
means a stronger test, as fewer networks are classified as software-like.

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be too low, because most software networks would be misclassified.
So we chose the value 0.88, that is immediatly greater than the second greater
non-software network S-score. With this value, we have both a high recall
(95.4\%) and a high precision (96.9\%).

%threshold 0.88: 2/66 non-sw misclassified, 3/65 sw misclassified
%precision: 62 / 64
%recall: 62 / 65

@@

We have chosen five different module dependency graphs, which where extracted
from actual dependencies between archives of five different software systems of
our sample: GEF (2 archives), iBATIS (4 archives), MegaMek (8 archives),
findbugs (16 archives), and zk (32 archives). 
%The module dependency networks are shown on Figure \ref{fig:architectures}. 
%\begin{figure*}[!t]
%\centering
%\includegraphics[width=1.0\textwidth]{architectures}
%\caption{Module dependency graphs}
%\label{fig:architectures}
%\end{figure*}

For the remaining parameters, the following values were given:

\begin{itemize}
\item $p, q, r \in {0.0, 0.2, 0.4, 0.6, 0.8, 1.0}$, with $p + q + r = 1$ and $p
+ q > 0$;
\item $din, dout \in {0, 1, 2, 3, 4}$;
\item $\mu in {0.0, 0.2, 0.4, 0.6}$.
\end{itemize}

We chose $p + q > 0$ because otherwise no node would be created by the model
after the creation of the initial network. We also avoided large values for
$\mu$ in order to ignore networks with highly-coupled modules.

In total, 9,500 BCR+ networks were synthesized.

The parameters values were chosen as such:

\begin{itemize}
\item $p1, p2, p3, p4 \in {0.0, 0.2, 0.4, 0.6, 0.8, 1.0}$, with $p1 > 0$ and $p1
  + p2 + p3 + p4 = 1$;
\item $e1, e2, e3, e4 \in {1, 2, 4, 8}$;
\item $\alpha \in {-1, 0, 1, 10, 100, 1000}$;
\item number of modules: $m \in {2, 4, 8, 16, 32}$.
\end{itemize}

Notice that $e_i$ has no effect when $p_i = 0$; in this case $e_i$ was just set
to zero.

In total, 38,790 CGW networks were synthesized.

@@

The following values were chosen:

\begin{itemize}
  \item $\mu in {0.0, 0.2, 0.4, 0.6}$;
  \item degree exponent distribution: $\gamma \in {2.18, 2.70, 3.35}$;
  \item module size distribution: $\beta \in {0.76, 0.99, 1.58}$;
  \item average degree: $k \in {5, 10, 15, 25}$;
  \item maximum degree: $max_k \in {58, 157, 482}$;
  \item minimum module size: $min_m \in {1, 10, 273}$.
\end{itemize}

In order to choose these values, we analyzed software networks from our sample
with approximately 500 to 2,000 nodes, so no network was much bigger or much
smaller than the synthetic networks. We computed the exponents for the degree
and module size distributions using the maximum likelihood estimation method
CITE-NEWMAN, and then chose the minimum, median and maximum values. For $k$,
$maxk$, and $min_m$, the values extracted from the software networks were
divided by the number of nodes in the network and then multiplied by 1000. We
then selected the minimum, median and maximum values. The parameter $max_m$ was
left unbound to avoid impossible combinations of parameters.

In total, 1,296 LF networks where synthesized.

@@

Each synthesized network was classified as software-like or non software-like,
using the classification model presented in section \ref{sec:classmodel}. The
results are summarized in Table \ref{tab:results}.

\begin{table}
\caption{Results for the classification of synthetic networks}
Table: Model | Networks classified as software-like
\begin{tabular}{ l | l }
\hline
Model & Networks classified as software-like \\
\hline \\
\hline
BCR+ & 21.18\% \\ % 2012 / 9500
CGW  & 19.40\% \\  % 7524 / 38790
LF   & 31.25\% \\ %  405 / 1296
\hline
\end{tabular}
\label{tab:results}
\end{table}

All models synthesized both software-like and non software-like networks. The
proportion of software-like networks was greater than 19\% for all models,
discarding the possibility that this result was obtained by pure chance.

%Show some graphs: histogram of average correlations for each model.
%\ref{fig:histograms}
%\begin{figure*}[!t]
%\center
%
%\subfigure[CGW]{\includegraphics[width=1.0in]{hist-cgw}}
%
%\subfigure[BCR+]{\includegraphics[width=1.0in]{hist-bcr}}
%
%\subfigure[LF]{\includegraphics[width=1.0in]{hist-lf}}
%
%\caption{Triad profiles}
%\label{fig:histograms}
%\end{figure*}

Of course, this result is of little practical value unless there is a
relationship between parameter values and S-score. For the purpose of this
research, it is important to know which values are more likely to lead to
software-like networks.

The algorithm 1R from data mining was used to help discover such relationship.
It analyzes the parameters and the classification of each network and finds a
rule that relates the value of one single parameter with the classification. We
are interested in rules for networks classified as software-like and in the
accuracy of these rules, i.e., the proportion of networks that are correctly
classified. The rules found by 1R are shown in Table \ref{tab:rules}.

\begin{table}
\caption{Rules for predicting the classification of a synthetic network. S
stands for software-like and N stands for non software like.}
Table: Model | Networks classified as software-like
\begin{tabular}{ l | l | l }
\hline
Model & Rule & Accuracy \\
\hline \\
\hline
\multirow{2}{*}{BCR+}
     & $\alpha \ge 0.7 \Rightarrow S$ & \multirow{2}{*}{82.4\%}  \\ 
     & $\alpha < 0.7 \Rightarrow N$ & \\ 
\hline
\multirow{2}{*}{CGW}
     & $p_1 \ge 0.5 \Rightarrow S$ & \multirow{2}{*}{82.3\%} \\  
     & $p_1 < 0.5 \Rightarrow N$ & \\  
\hline
\multirow{2}{*}{LF}   
     & $\gamma \ge 2.44$ & \multirow{2}{*}{78.9\%} \\ 
     & $\gamma \ge 2.44$ &  \\ 
\hline
\end{tabular}
\label{tab:rules}
\end{table}

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% in all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.

%\subsection{Parameter-Based Classification Models}
%
%1R

%Naive Bayes

%\subsection{Homogeinity}
%
%Pick realistic networks from a model. Are they similar to each other? (see
%standard deviation) In other words: do the parameters make a difference?
%
%Are they similar to networks generated by the other models? In other words: are
%the models equivalent?

%\section{Threats to validity}
%
%% External validity (EV): can it be generalized?
%
%%Structural information isn't enough for a expert do produce a
%%decomposition (s/he may use data such as names and external documentation). 
%%
%%Even when considering only structural information, is it true
%%that experts would find a decomposition similar to the reference decomposition
%%imposed by the model?
%
%We generated only one network for each set of parameters. (as we've shown, some
%parameters are redundant as they do not change significantly the realism)
%
%Some clustering algorithms use weights and they weren't studied here.
%
%EV: We've only studied 65 systems, which is not that much.
%
%We only studied object-oriented systems implemented in Java. Maybe the results
%would be different if we studied systems implement in other languagens or using
%other paradigms. The choice of a particular technique for extracting
%dependencies (static analysis) may also have impact on the structure of the
%networks.
%
%% Koschke says that experts decompositions vary by no more than 80? percent.

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously studied with real networks
\cite{Wu2005}. After that we will be able to compare the results obtained by
the approaches.

\appendix{Network Data Set}

\section{Software Networks}
